\chapter{CONCLUSIONS}

This project successfully developed FIRMAI, an integrated AI-powered security framework for comprehensive IoT vulnerability analysis combining firmware emulation, network scanning, and intelligent malware detection. We achieved all primary objectives establishing unified platform eliminating fragmentation, implementing high-efficiency firmware emulation with 78.4\% success rate approaching 5$\times$ improvement over baseline 16\%, creating comprehensive network discovery completing Class C scans in under 5 minutes with 88.5\% device classification accuracy, and deploying AI-powered malware detection achieving 99.28\% classification accuracy with GPU-accelerated real-time processing capability.

\section{Future Enhancements}

Several promising directions exist for extending FIRMAI capabilities. Enhanced firmware analysis could incorporate automated exploit generation, machine learning-based vulnerability prediction, symbolic execution integration, and firmware patching automation. Advanced network capabilities could add support for additional protocols including BLE, Zigbee, Z-Wave, enhanced topology inference using machine learning, anomaly detection for unusual behaviors, and SDN/NFV integration. Malware detection improvements could implement online learning, adversarial robustness training, explainable AI, and transfer learning. System scalability enhancements could deploy distributed architecture, implement horizontal scaling with Kubernetes, add GPU cluster support, and optimize resource allocation. User experience improvements could develop mobile applications, implement advanced visualization, add natural language query interface, and create automated remediation workflows.

\section{Limitations}

Despite significant achievements, limitations exist. Firmware emulation limitations include 78.4\% success rate meaning 21.6\% cannot be emulated, limited PowerPC support, timeout constraints, and NVRAM emulation imperfections. Network scanning limitations include primarily IPv4 focus, Layer 2 ARP requiring same subnet, firewall interference, and passive discovery absence. Malware classification limitations include training data dependency, novel threat challenges, false positive rate of 0.28\%, and protocol-specific focus. Integration limitations include manual configuration required, single-server deployment in current implementation, limited SIEM integration, and customization complexity. Performance limitations include GPU dependence, resource requirements, parallel scaling limits, and database optimization opportunities.

\section{Final Remarks}

The development of FIRMAI represents significant advancement in IoT security addressing critical gaps through intelligent integration of multiple security domains. The 78.4\% firmware emulation success rate demonstrates substantial improvement while 99.28\% malware classification accuracy validates deep learning effectiveness. Our open-source approach ensures accessibility enabling organizations with limited budgets to deploy professional-grade security capabilities. The modular architecture provides foundation for continued enhancement ensuring long-term relevance. We believe FIRMAI makes meaningful contribution to IoT security landscape providing practical tools while advancing state-of-the-art in automated security analysis.

% Bibliography
\begin{thebibliography}{99}
	
	\bibitem{chen2016}
	Chen, D. D., Woo, M., Brumley, D., \& Egele, M. (2016). Towards automated dynamic analysis for Linux-based embedded firmware. \textit{NDSS}.
	
	\bibitem{kim2020}
	Kim, M., Kim, D., Kim, E., Kim, S., Jang, Y., \& Kim, Y. (2020). FirmAE: Towards Large-Scale Emulation of IoT Firmware for Dynamic Analysis. In \textit{Annual Computer Security Applications Conference}.
	
	\bibitem{khan2023}
	Khan, I. A., Moustafa, N., Pi, D., Haider, W., Li, B., \& Jolfaei, A. (2023). An enhanced multi-stage deep learning framework for detecting malicious activities from autonomous vehicles. \textit{IEEE Transactions on Intelligent Transportation Systems}.
	
	\bibitem{abbas2023}
	Al Abbas, M., Al-Hawawreh, M., \& Aljuhani, A. (2023). Intelligent IoT malware detection using RNN-based NLP techniques. In \textit{2023 International Conference on Cyber Security and Internet of Things}.
	
	\bibitem{jeyalakshmi2024}
	Jeyalakshmi, S., \& Nallaperumal, K. (2024). Hybrid CNN-GAN and DBN Based IoT Malware Detection. \textit{Journal of Network Security}.
	
	\bibitem{chaganti2022}
	Chaganti, R., Ravi, V., \& Pham, T. D. (2022). Cross-architecture deep learning based Android malware detection. \textit{Journal of Information Security and Applications}.
	
	\bibitem{stratosphere}
	Stratosphere Laboratory. (2020). IoT-23 Dataset. Available at: \url{https://www.stratosphereips.org/datasets-iot23}
	
	\bibitem{nmap}
	Lyon, G. F. (2009). \textit{Nmap Network Scanning: The Official Nmap Project Guide to Network Discovery and Security Scanning}. Insecure.
	
\end{thebibliography}