\chapter*{ACKNOWLEDGEMENTS}
	\addcontentsline{toc}{chapter}{Acknowledgements}


	We express our deep sense of gratitude to all who helped us complete this project successfully. We are deeply indebted to our Project Supervisor \textbf{Ms. Praseetha S Nair} for her excellent guidance and valuable comments throughout the project duration. We express sincere gratitude to our project coordinator \textbf{Ms. Salitha M.K} for her timely guidelines and critical reviews that shaped this work.
	
	We are thankful to our Head of Department \textbf{Ms. Salitha M.K} for her continuous support and encouragement. We also thank all faculty members of the Department of Internet of Things \& Cyber Security who helped us during this project work through their valuable suggestions and technical discussions.
	
	We extend special thanks to the global research community whose published work in firmware emulation, network security, and machine learning guided our implementation approach. The foundational research in these areas provided invaluable insights that informed our system design and development methodology.
	
	Finally, we thank our parents, family members and friends who directly and indirectly contributed to the successful completion of our project through their constant support and encouragement.
	
	\vspace{1cm}
	
	\begin{tabular}{ll}
		\textbf{ABHIRAM G NAIR} & PRC22CSOT001 \\
		\textbf{NAVOMI TITUS} & PRC22CSOT019 \\
		\textbf{SPANDANA NAIR} & PRC22CSOT025 \\
		\textbf{VIGNESH S KUMAR} & PRC22CSOT028 \\
	\end{tabular}
	
	\vspace{0.5cm}
	
	\noindent\textbf{Date:} 19/03/2026
	
	% Abstract